Integrating Big Data and Deep Learning for Improved Wheat (Triticum aestivum) Grain Yield Predictions

### 2023-05-24 submission_informatika.tex
shortned the text

\editor{Abhi} % Names of Editors

@@

\begin{itemize}
    \item To propose a framework that effectively combines genomic, environmental, and GxE input \
data for utilization in CNNs.
    \item To benchmark the developed framework against Extended GBLUP models using cross-validation\ 
scenarios.
\end{itemize}

@@

where \(y\) is the observation corrected for environmental effect (trial) for grain yield, \(G\) is the design matrix of genotypes, \(E\) is the design matrix of environments, \(\tau\) is the vector of genotypic effects, \(u\) is the vector of environmental effects, and \(e\) is the vector of residuals.

@@

\textit{Integration of phenotypic, genotypic, and environment data was done to prepare the data for benchmarking with cross-validation scenarios.}

Two broad cases for grain yield prediction were considered. Firstly, prediction of mean genotypic values across environments, and secondly, prediction of mean genotypic values within environments. For the first case, only genotypic data was used as an input, whereas for the second case, genotypic data along with environment data was used as an input. Additionally, GxE data was approximated following \cite{de2020data}.

@@

For the second case, i.e., prediction of mean genotypic values within environments, E-GBLUP was extended (referred to as Ex-GBLUP) to include environmental and GxE effects in addition to the effects of genotypes (Tab. \ref{exGBLUP_models}). An all-CNN model was developed with multiple input streams feeding into a concatenation layer, followed by dense layers. Each stream corresponds to different model inputs, namely genomic data plus twenty-eight weather variables (Fig. 2). The hyperparameters of the all-CNN model were tuned using the hyperband optimization algorithm \cite{li_hyperband_2018}. For all-CNN, a validation set was additionally created by subsetting 20\% from the respective train and test sets. Four cross\-validation scenarios were evaluated: (1) CV1 - predicting known genotypes in known environments, (2) CV2 - predicting known genotypes in new environments, (3) CV3 - predicting new genotypes in known environments, and (4) CV4 - predicting new genotypes in new environments.

\begin{table}[htbp]
\centering
\setlength{\tabcolsep}{5pt}
\begin{tabular}{ l c c c c c c c c  l }
 \hline
 Model & E{\textunderscore}I & G{\textunderscore}I & G{\textunderscore}a & G{\textunderscore}d & G{\textunderscore}aa & $\textit{ERM}_l$ & $G_a \otimes \textit{ERM}_l$  \\
 \hline
 M\textunderscore1 & 1 & 1 &   &   &   &    &    \\
 M\textunderscore2 & 1 &   & 1 & 1 &   &    &    \\
 M\textunderscore3 & 1 &   & 1 & 1 & 1 &    &    \\
 M\textunderscore4 &   &   & 1 & 1 & 1 & 1  &    \\
 M\textunderscore5 &   &   & 1 & 1 & 1 & 1  & 1  \\
 \hline
\end{tabular}
\caption{Description of the tested models for the prediction of the mean values of the individual sites. The entry "1" means that the respective data was considered in the model. $G_a \otimes \textit{ERM}_l$ is the element wise product of respective matrices. $E_I$ and $G_I$ are identity matrices for environments and genotypes. The matrices $G_a$, $G_d$, and $G_{aa}$ are relationship matrices of genotypes used for additive marker effects, dominance effects, and additive times additive epistasis effects, respectively.}
\label{exGBLUP_models}
\end{table}

@@

@@

\textit{All-CNN benefits from increased training set size but fails to outperform Ex-GBLUP}

The results from 25 cross-validation runs per combination of nine models and four scenarios are summarized in Figure \ref{fig_for_informatika_cv}. The findings indicate that models incorporating genomic and environmental relatedness consistently outperform those that ignore relatedness, except for CV2. Notably, the prediction accuracy for non-hybrid performance is consistently higher than that for hybrid performance across all scenarios. Complex models demonstrate superior performance in the CV3 and CV4 validation scenarios compared to other models. Specifically, model M5 and M6 exhibit similar performance trends, outperforming M4 in CV3 and CV4, while not showing the same advantage in CV1 and CV2. These results emphasize the importance of considering genomic and environmental relatedness in predictive.

\begin{figure}[h]
    \centering
    \includegraphics[width=.8\textwidth]{for_informatika_cv}
    \caption{Genomic prediction accuracy for mean values (mean{\textunderscore}v) of grain yield of individual locations for the validation scenarios CV1, CV2, CV3 and CV4.}
    \label{fig_for_informatika_cv}
\end{figure}

>>>>

\editor{Abhishek} % Names of Editors

@@

\begin{itemize}
    \item Propose and discuss a framework to integrate genotypic, phenotypic and environment data for use in predicting wheat grain yield
    \item Develop and benchmark a CNN model against extended G-BLUP (EG-BLUP)
\end{itemize}

@@

where \(y\) is vector of the design effect corrected grain yield values, \(G\) is the design matrix of genotypes, \(E\) is the design matrix of environments, \(\tau\) is the vector of genotypic effects, \(u\) is the vector of environmental effects, and \(e\) is the vector of residuals.

@@

@@

@@

\textit{The environment relationship matrix reveals no environment relationship structure}

The ERM was utilized to examine the presence of environmental clusters, but the analysis did not uncover any discernible pattern. This finding can be interpreted in two ways: 1. There are no environmental clusters present in the tested environments, or 2. There was an error in the calculation of the ERM. The latter possibility could be attributed to either poor-quality input data or an incorrect calculation method. However, since the data is sourced from a reliable pipeline, its quality is assured. To further investigate the calculation method, the ERM was computed using intervals of 5 and 10 days, in addition to the original 30-day interval. The null hypothesis, suggesting no relationship between ECs derived from the 5, 10, and 30-day intervals, was rejected based on the results of the mantel.test() function in R (Paradis and Schliep, 2019). Consequently, it was concluded that the examined environments lack any underlying structure and are truly diverse. Studies have demonstrated that environment relationship matrix can be used to derive an approximated GxE matrix \cite{de2020data}, but the same was not a part of curent study.

@@
